Experiences with Planning for Natural Language Generation

### 2009-05-04 trunk/Papers/ci-spark/manuscript.tex
gone over domains chapter

This view of generation as planning has a long tradition in NLG
\citep{perrault80,appelt:planning,hovy88,young94dpocl}. Despite their
theoretical attractiveness, these early approaches never gained
mainstream use in applied systems, in part because of the inefficiency
of the planners that were available at the time. However, the recent
improvements in planner efficiency and expressiveness has sparked a
renewed interest in applying modern planning techniques to NLG---both
on the level of speech act planning
\citep{Steedman-Petrick:07,benotti08b} and and on the level of
sentence generation \citep{KolSto07}. The goal of this paper is to
determine, by looking at some representative generation problems,
whether planning has advanced to the point that this enterprise can be
successful this time.

@@

Problems in NLG at all levels of description can become
computationally complex very easily. For instance, \citet{KolStr02}b
showed that even surface realization is NP-complete under reasonable
assumptions. This makes it necessary to investigate methods for
guiding the search, which makes modern planning techniques a promising
approach to NLG. The point of this paper is to investigate whether
existing planners fulfill this promise of efficiency for the
application to NLG.

@@

Sentence generation is the problem of computing from a grammar and a
semantic representation, a single sentence that expresses this piece of
meaning. Traditionally, this problem is split into several steps
\citep{reiter00building}. In a first step, called \emph{sentence planning},
the semantic representation is enriched with additional information. For
instance, \emph{referring expressions}, which refer to individuals (e.g.,
objects, people, etc.) in the domain of discourse, are determined at this
point. In a second step, called \emph{surface realization}, this enriched
representation is translated into a natural-language sentence, using the
grammar.

In practice, the process of determining referring expressions typically
interacts with the realization step and so it often turns out to be
beneficial to perform these two steps together. For instance, the SPUD
system (``Sentence Planning Using Description''; \citealt{Stone2003a})
performed this task using a top-down generation algorithm based on
tree-adjoining grammars \citep{joshi;etal1997} whose lexical entries were
equipped with semantic and pragmatic information. Unfortunately, SPUD had
to resort to a non-optimal greedy search strategy to cope with its large
search space and retain reasonable efficiency. To improve the efficiency
of SPUD, \citet{KolSto07} translate the sentence generation problem into a
planning problem and use a planning algorithm to perform the generation
task.

@@

We illustrate this process on a simplified example.  Consider a
knowledge base containing the individuals $e$, $r_1$ and $r_2$, and a
set of attributes encoding the fact that $r_1$ and $r_2$ are rabbits,
$r_1$ is white and $r_2$ is brown, and $e$ is an event in which $r_1$
sleeps.  Say that we want to express the information
$\{\mathsf{sleep}(e,r_1)\}$ using the tree-adjoining grammar shown in
Figure~\ref{fig:white-rabbit-sleeps-grammar}. This grammar consists of
\emph{elementary trees} (i.e., the disjoint trees in the figure), each
of which contributes certain \emph{semantic content}. We can
instantiate these trees by substituting individuals for \emph{semantic
  roles}, such as $\mathsf{self}$ and $\mathsf{subj}$, and then
combine the tree instances as shown in
Figure~\ref{fig:white-rabbit-sleeps-deriv} to obtain the sentence
``The white rabbit sleeps''.

We compute this grammatical derivation in a top-down manner, starting with
the elementary tree for ``sleeps''. This tree satisfies the need to convey
the semantic information, but introduces a need to generate a noun phrase
(NP) for the subject; this NP must refer uniquely to the target referent
$r_1$. In a second step, we substitute the tree for ``the rabbit'' into the
open NP leaf, which makes the derivation grammatically complete. Since
there are two different individuals that could be described as ``the
rabbit''---technically, $r_2$ is still a \emph{distractor} (i.e., based on
the description ``the rabbit'', the hearer might erroneously think that
we're talking about $r_2$ and not $r_1$)---we are still not finished. To
complete the derivation, the tree for ``white'' is added to the existing
structure by an \emph{adjunction} operation, making the derivation
both syntactically and semantically complete.

@@

The process described above has clear parallels to planning: we manipulate
states by applying sequences of actions to achieve a goal. We can make this
connection even more precise by translating the SPUD problem into a
planning problem.\footnote{See
  \url{http://code.google.com/p/crisp-nlg/} for the CRISP system, in
  which this conversion is implemented.}  For instance,
Figure~\ref{fig:white-rabbit-as-planning} shows the corresponding PDDL
actions for the above generation task, where each action corresponds
to an operation that adds a single elementary tree to the derivation.
In each case, the first parameter of the action is a node name in the
derivation tree, and the remaining parameters stand for the
individuals to which the semantic roles will be instantiated.  The
syntactic preconditions and effects are encoded using $\mathsf{open}$
and $\mathsf{canadjoin}$ literals; the status of each referring expression
is tracked using $\mathsf{distractor}$ literals. The resulting action
effects contain terms of the form $\mathsf{subj}(u)$, which construct new
node names, but aren't strictly permitted by PDDL. In order to meet the
syntactic requirements of PDDL, these terms can be eliminated by estimating
an upper bound $n$ for the plan length, making $n$ copies of each action,
ensuring that copy $i$ can only be applied in step $i$, and replacing the
term $\mathsf{subj}(u)$ in an action copy by the constant
$\mathsf{subj}_i$.

Once the appropriate actions are defined, we can solve the generation
problem as an ordinary planning problem. For instance, the following plan
solves our previous example:
%
\begin{enumerate}
\item $\mathsf{add}\textsf{-}\mathsf{sleeps}(\mathsf{root}, r_1)$,
\item $\mathsf{add}\textsf{-}\mathsf{rabbit}(\mathsf{subj}(\mathsf{root}),r_1)$,
\item $\mathsf{add}\textsf{-}\mathsf{white}(\mathsf{subj}(\mathsf{root}),r_1)$.
\end{enumerate}
%
Using this plan, the grammatical derivation in
Figure~\ref{fig:white-rabbit-sleeps-deriv}, and therefore the
generated sentence ``the white rabbit sleeps'', can be systematically
reconstructed. Thus, we can solve the sentence generation problem via
the detour through planning and bring current search heuristics for
planning to bear on generation.

@@

The object of the GIVE Challenge (``Generating Instructions in Virtual
Environments''; \citealt{alexander07:_shared_task_propos}) is to build an
NLG system which is able to produce natural-language instructions which
will guide a human user in performing some task in a virtual environment.
From an NLG perspective, GIVE makes for an interesting challenge because it
is a theory-neutral task that exercises all components of an NLG system,
and emphasizes the study of communication in a (simulated) physical
environment. It also has the advantage that the user and the NLG system can
be in physically different places, as long as the 3D client and the NLG
system are connected over a network. This makes it possible to evaluate
GIVE NLG systems on a large scale over the Internet.
%
The first GIVE evaluation is currently underway. Five NLG systems from
three countries are in the process of being evaluated. At the time of
writing, 1028 different evaluation runs---each consisting of one user
interacting with one instance of the virtual environment and being guided
by an NLG system---have been collected, making GIVE the largest-scale
NLG evaluation effort ever.\footnote{See the GIVE website at
 \url{http://www.give-challenge.org/} for more details about this
 challenge.}

\begin{figure}[t]
\centering
%\includegraphics[width=1 \columnwidth]{give_world_no_expl}
\includegraphics[width=0.75\columnwidth]{give_world_2}
\caption{Map of the GIVE development world.}
  \label{fig:give-development-world}
\end{figure}

An example GIVE world map is shown in
Figure~\ref{fig:give-development-world}. To simplify both the planning and
the NLG task, the world is discretised into a set of tiles of equal size.
The user can turn by 90 degree steps in either direction, and can move from
the centre of one tile to the centre of the next tile, provided the path
between two tiles is not blocked. The world also consists of a set of
objects which can be manipulated by the user in various ways. For
instance, in the example world the user's task is to pick up a trophy in
the top left room. The trophy is hidden in a safe behind a picture, so the
user must first move the picture out of the way and open the safe by
pushing a sequence of buttons (the small square boxes on the walls) in the
correct order. In order to get to all these buttons, the user must also
open the door in the centre of the map and deactivate the alarm tile by
pushing further buttons. Figure~\ref{fig:give-planning} shows the encoding
of some of the available GIVE domain actions, in PDDL syntax.

@@

(:action manipulate-b1-off-on
   :parameters (?pos - position)
   :precondition 
       (and (state b1 off)
            (player-pos ?pos)
            (position b1 ?pos))
   :effect 
       (and (not (state b1 off))
            (state b1 on)
            (not (state d1 closed))
            (state d1 open) 
            (not (blocked pos_6_5 pos_6_4))
            (not (blocked pos_6_4 pos_6_5))))
\end{verbatim}}%
\end{minipage}
\caption{PDDL actions for the GIVE domain.}
\label{fig:give-planning}
\end{figure}

The plans generated in the GIVE domain are often nontrivial. For instance,
in the above example the shortest plan consists of 108 action steps, with
the first few steps as follows:
%
\begin{enumerate}
\item $\mathsf{turn}\textsf{-}\mathsf{left}(\mathsf{north},
\mathsf{west})$,
\item $\mathsf{move}(\mathsf{pos\_5\_2}, \mathsf{pos\_4\_2}, \mathsf{west})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{b1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_5\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{west}, \mathsf{north})$.
\end{enumerate}

Most of the actions in the above plan are ``move'' actions. This makes the
bulk of the GIVE problem very similar to that of a Gridworld problem (see,
e.g., \citep{Tovey-Koenig:2000}), which also involves finding a route
through a two-dimensional world map with discrete positions. GIVE tasks are
usually much more complex, however, as illustrated by the above example
map: a successful plan must include steps for pressing buttons in the right
order, reasoning about large numbers of world objects, and navigating
through complicated room shapes.

The GIVE task comes with a number of interesting research challenges for
both the NLG and planning communities. On the NLG side, one major problem
is that verbalising each action instance in a plan makes for very
poor natural-language instructions. For instance, the NLG system could
generate an instruction sequence starting with ``turn left; walk forward;
press the button; turn right; walk forward; walk forward; turn right; walk
forward; walk forward; turn left; walk forward''. However, directions of
this form are linguistically clumsy and uninteresting, and it is to be
expected that frustrated users will quickly cancel an interaction with such
a system, resulting in a bad evaluation score. Having just pressed the
button, it would be much better to say ``turn around and walk through the
door''. That is, the system should \emph{summarise} plans by merging
multiple action instances into single instructions before presenting them
to the user.\footnote{This task is similar to the construction of
  ``macro'' actions, but is primarily intended for plan presentation
  purposes. See, e.g., \citep{Botea-etal:05} which contains a survey of
  recent approaches for generating macro actions in planning.}
Conversely, it may also be necessary to \emph{elaborate} on a single
planning step by expressing it with several instructions. Having just
entered the top left room, it may be easier for the user to understand the
instruction ``walk to the centre of the room; turn right; now press the
green button in front of you'', rather than the instruction ``press the
green button on the wall to your right''. Furthermore, in order to plan the
referring expression ``the green button in front of you'' at a time when
the button is not yet in front of the user, the NLG system must keep track
of the hypothetical changes in what objects are visible to the user. Thus,
the NLG acts of referring and instructing must be tightly integrated with
the modules for plan generation.

From a planning perspective, GIVE imposes very strict runtime
requirements on the plan generation task: planning must
happen in real time and the system must respond to a user in a timely
fashion.  If the system takes too much time deliberating over an
instruction to give, rather than actually giving this instruction, the
user may have walked or turned away, thus making the instruction
invalid.  Furthermore, plan execution monitoring also plays an
important role in the GIVE problem.  At a high level, the system needs
to monitor a user's actions and compare them against the generated
instruction set to determine if the user has correctly followed
directions or not. In the case of the latter, new instructions may
have to be generated. In practice, the situation can be quite
complicated since the mental state of the user is not known and so the
system must be inferred from observing the user's actions in real
time. For example, a user directed to ``turn around and walk through
the door'' may not necessarily perform these actions to the letter,
i.e., immediately turning 180 degrees and proceeding directly to the
door. Instead, the user might take a roundabout route through the
room, eventually exiting out the door.  Although the user's actions do
not match the generated instructions exactly, they meet the intended
goal. The system must be able to identify such ``equivalent'' plans
and not immediately generate new instructions as soon as the user's
actions have gone off course. Furthermore, a user can communicate
certain intentions to the system, both through action and
inaction. For instance, the system should infer that a user has failed
to follow instructions if the user exits a room when given a directive
to ``walk to the centre of the room''. The system should also make a
similar conclusion if a user simply does nothing when given the
instruction.

@@

>>>>

This view of generation as planning has a long tradition in NLG
\citep{perrault80,appelt:planning,hovy88,young94dpocl}. Despite their
theoretical attractiveness, these early approaches never gained
mainstream use in applied systems, in part because of the inefficiency
of the planners that were available at the time. However, the recent
improvements in planner efficiency and expressiveness has sparked a
renewed interest in applying modern planning techniques to NLG---both
on the level of speech act planning
\citep{Steedman-Petrick:07,benotti08b} and and on the level of
sentence generation \citep{KolSto07}. The goal of this paper is to
determine, by looking at some representative generation problems,
whether planning has advanced to the point that this enterprise can be
successful this time---and thus, to determine the usefulness of
current planning technology to an application area that has a long
tradition, but is not currently in the focus of the planning
competitions.

@@

Problems in NLG at all levels of description can become
computationally complex very easily. For instance, \citet{KolStr02}
showed that even surface realization is NP-complete under reasonable
assumptions. This makes it necessary to investigate methods for
guiding the search, which makes modern planning techniques a promising
approach to NLG. The point of this paper is to investigate whether
existing planners fulfill this promise of efficiency for the
application to NLG.

@@

Although the sentence generation task is often split up into two
separate steps for sentence planning and surface realization, the
decisions that an NLG system must make in these two modules often
interact. For instance, in order to generate \emph{referring
  expressions (REs)}, i.e.\ unique descriptions of individuals in the
universe of discourse, it is necessary to know what properties of
individuals can be expressed as nouns and adjectives. Grammatical
information like this should belong exclusively to the surface
realization module; but on the other hand, RE generation enriches the
input meaning representation and is therefore usually seen as part of
sentence planning.

One seminal approach to solving the sentence planning and surface
realization problems together is offered by the SPUD generation system
(``Sentence Planning Using Description''; \citealt{Stone2003a}). This
approach assumes a tree-adjoining grammar (TAG) \citep{joshi;etal1997}
whose lexical entries are equipped with semantic and pragmatic
information, as well as two sets of ground atoms: the
\emph{communicative goal} that specifies what semantic information we
must express, and a \emph{knowledge base}. Let's consider an example
to illustrate the problem that SPUD solves. Consider a knowledge base
containing the individuals $e$, $r_1$ and $r_2$, and a set of
attributes encoding the fact that $r_1$ and $r_2$ are rabbits, $r_1$
is white and $r_2$ is brown, and $e$ is an event in which $r_1$
sleeps.  Say that we want to express the information
$\{\mathsf{sleep}(e,r_1)\}$ using the tree-adjoining grammar shown in
Figure~\ref{fig:white-rabbit-sleeps-grammar}. This grammar consists of
\emph{elementary trees} (i.e., the disjoint trees in the figure), each
of which contributes certain \emph{semantic content}. We can
instantiate these trees by substituting individuals for \emph{semantic
  roles}, such as $\mathsf{self}$ and $\mathsf{subj}$, and then
combine the tree instances as shown in
Figure~\ref{fig:white-rabbit-sleeps-deriv} to obtain the sentence
``The white rabbit sleeps''.

@@

We can now consider various different algorithms for solving the SPUD
problem. The naive algorithm builds the TAG derivation top-down. In
the example, it starts with the elementary tree for ``sleeps''. This
tree satisfies the need to convey the semantic information, but
introduces a need to generate a noun phrase (NP) for the subject,
which must refer uniquely to the target referent $r_1$. In a second
step, we substitute the tree for ``the rabbit'' into the open NP leaf,
which makes the derivation grammatically complete. Since there are two
different individuals that could be described as ``the
rabbit''---technically, $r_2$ is still a \emph{distractor} (i.e.,
based on the description ``the rabbit'', the hearer might erroneously
think that we're talking about $r_2$ and not $r_1$)---we are still not
finished. To complete the derivation, the tree for ``white'' is added
to the existing structure by an \emph{adjunction} operation, making
the derivation both syntactically and semantically complete. This
search algorithm takes worst-case exponential time because it may have
to consider many combinations of elementary trees; and in fact, the
SPUD problem generalizes the surface realization problem considered by
\citet{KolStr02} and is therefore NP-complete.

@@

The SPUD system itself worked around this complexity by using a
greedy, incomplete search algorithm. On the other hand,
\citet{KolSto07} present an alternative to controlling the necessary
search, by encoding the generation problem into a planning
problem.\footnote{See \url{http://code.google.com/p/crisp-nlg/} for
  the CRISP system, in which this conversion is implemented.} Each
operator of this planning problem models the addition of some
elementary tree to the TAG derivation. For instance,
Figure~\ref{fig:white-rabbit-as-planning} shows the corresponding PDDL
actions for the above generation task. The \texttt{add-sleeps}
operator represents the addition of the elementary tree for ``sleeps''
to the derivation. $\verb?add-sleep?(\mathsf{root},e,r_1)$ removes the
atom $\mathsf{subst}(S,\mathsf{root})$, which expresses that there is
currently an open substitution node $\mathsf{root}$ with label $S$,
from the planning state and replaces it with an atom
$\mathsf{subst}(NP,\mathsf{subj}(\mathsf{root}))$; here
$\mathsf{subj}(\mathsf{root})$ is meant as shorthand for a fresh
individual name.\footnote{These terms, which are not valid in ordinary
  PDDL, can be eliminated by estimating an upper bound $n$ for the
  plan length, making $n$ copies of each action, ensuring that copy
  $i$ can only be applied in step $i$, and replacing the term
  $\mathsf{subj}(u)$ in an action copy by the constant
  $\mathsf{subj}_i$. Notice that $S$, $NP$, and $N$ are constants.}
At the same time, the operator records that the semantic information
$\mathsf{sleep}(e,r_1)$ has now been expressed, and introduces all
individuals except for $r_1$ as distractors for the new RE at
$\mathsf{subj}(\mathsf{root})$. These distractors can then be removed
by subsequent applications of the other two operators. By assuming an
initial state that encodes the knowledge base and contains an atom
$\mathsf{subst}(S,\emph{root})$, and a goal that contains, among
others, $\forall x \forall y. \neg \mathsf{subst}(x,y)$ and
$\mathsf{expressed}(\emph{sleep},e,r_1)$, a planner will find plans
like the following:
%
\begin{enumerate}
\item $\mathsf{add}\textsf{-}\mathsf{sleeps}(\mathsf{root}, r_1)$,
\item $\mathsf{add}\textsf{-}\mathsf{rabbit}(\mathsf{subj}(\mathsf{root}),r_1)$,
\item $\mathsf{add}\textsf{-}\mathsf{white}(\mathsf{subj}(\mathsf{root}),r_1)$.
\end{enumerate}
%
Using this plan, the grammatical derivation in
Figure~\ref{fig:white-rabbit-sleeps-deriv}, and therefore the
generated sentence ``the white rabbit sleeps'', can be systematically
reconstructed. Thus, we can solve the sentence generation problem via
the detour through planning and bring current search heuristics for
planning to bear on generation.

@@

Let's now turn to a second recent application of planning in NLG, the
GIVE Challenge (``Generating Instructions in Virtual Environments'';
\citealt{ByrKolStrCasDalMooObe09}). The object of this shared task is
to build an NLG system which produces natural-language instructions
which will guide a human user in performing some task in a virtual
environment.  From an NLG perspective, GIVE makes for an interesting
challenge because it is a theory-neutral task that exercises all
components of an NLG system, and emphasizes the study of communication
in a (simulated) physical environment. Another advantage is that GIVE
makes it possible to connect NLG systems to users over the Internet,
and thus offer cheap access to human experimental subjects. The first
installment of GIVE (GIVE-1) evaluated five NLG systems on the
performance of 1143 users, making it the largest NLG evaluation effort
in terms of human users ever.

\begin{figure}[t]
\centering
%\includegraphics[width=1 \columnwidth]{give_world_no_expl}
\includegraphics[width=0.75\columnwidth]{give_world_2}
\caption{Map of a GIVE world.}
  \label{fig:give-development-world}
\end{figure}

Planning plays a central role in the GIVE task. To see this, consider
the example GIVE world map in
Figure~\ref{fig:give-development-world}. To simplify both the planning
and the NLG task, this world is discretised into a set of tiles of
equal size.  The user can turn by 90 degree steps in either direction,
and can move from the centre of one tile to the centre of the next
tile, provided the path between two tiles is not blocked. The world
also consists of a set of objects which can be manipulated by the user
in various ways. For instance, in the example world the user's task is
to pick up a trophy in the top left room. The trophy is hidden in a
safe behind a picture; to access it, the user must push certain
buttons in order to move the picture out of the way, open the safe,
and open doors. Figure~\ref{fig:give-planning} shows the encoding of
some of the available GIVE domain actions, in PDDL syntax. In the
example, the shortest plan to solve the task consists of 108 action
steps, with the first few steps as follows:
%
\begin{enumerate}
\item $\mathsf{turn}\textsf{-}\mathsf{left}(\mathsf{north},
\mathsf{west})$,
\item $\mathsf{move}(\mathsf{pos\_5\_2}, \mathsf{pos\_4\_2}, \mathsf{west})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{b1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_5\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{west}, \mathsf{north})$.
\end{enumerate}

@@

(:action manipulate-b1-off-on
   :parameters (?pos - position)
   :precondition 
       (and (state b1 off)
            (player-pos ?pos)
            (position b1 ?pos))
   :effect 
       (and (not (state b1 off))
            (state b1 on)
            (not (state d1 closed))
            (state d1 open) 
            (not (blocked pos_6_5 pos_6_4))
            (not (blocked pos_6_4 pos_6_5))))
\end{verbatim}}%
\end{minipage}
\caption{Some PDDL actions for the GIVE domain.}
\label{fig:give-planning}
\end{figure}

A GIVE NLG system must be able to compute such plans. At a minimum,
the discourse planner will call a planner in order to determine the
content of the instructions that should be presented to the
user. Under this view, the planning problem is very similar to the
Gridworld problem (see, e.g., \citep{Tovey-Koenig:2000}), which also
involves finding a route through a two-dimensional world map with
discrete positions. GIVE tasks are usually much more complex, however,
as illustrated by the above example map: a successful plan must
include steps for pressing buttons in the right order, reasoning about
large numbers of world objects, and navigating through complicated
room shapes. This relatively loose integration of NLG system and
planner is the state of the art of the systems that participated in
GIVE-1.

However, in order to solve the GIVE task in a more satisfying way, we
expect that planning and generation will have to be interleaved more
closely, making the planning a genuine part of the NLG task. For
instance, the NLG system could generate an instruction sequence
starting with ``turn left; walk forward; press the button; turn right;
walk forward; walk forward; turn right; walk forward; walk forward;
turn left; walk forward''. But such instructions are clumsy and
uninteresting; it would be much better to say ``turn around and walk
through the door''. That is, the system should \emph{summarise} plans
by merging multiple action instances into single instructions before
presenting them to the user.\footnote{This task is similar to the
  construction of ``macro'' actions, but is primarily intended for
  plan presentation purposes. See, e.g., \citep{Botea-etal:05} which
  contains a survey of recent approaches for generating macro actions
  in planning.}  Conversely, it may also be necessary to
\emph{elaborate} on a single planning step by expressing it with
several instructions. Having just entered the top left room, it may be
easier for the user to understand the instruction ``walk to the centre
of the room; turn right; now press the green button in front of you'',
rather than the instruction ``press the green button on the wall to
your right''. Furthermore, in order to plan the referring expression
``the green button in front of you'' at a time when the button is not
yet in front of the user, the NLG system must keep track of the
hypothetical changes in what objects are visible to the user. Thus,
the NLG acts of referring and instructing must be tightly integrated
with the modules for plan generation. This makes the GIVE domain
planning problem an integral part of the NLG system, which makes this
a second planning problem of immediate relevance to NLG that is
completely different from the first one.

% It seems to me that this paragraph misses the point. We're not
% trying to sell GIVE as a cool problem here; we take the perspective
% that GIVE exists and has already been shown to be cool, and we only
% talk about solving it here. - AK
%
% From a planning perspective, GIVE imposes very strict runtime
% requirements on the plan generation task: planning must
% happen in real time and the system must respond to a user in a timely
% fashion.  If the system takes too much time deliberating over an
% instruction to give, rather than actually giving this instruction, the
% user may have walked or turned away, thus making the instruction
% invalid.  Furthermore, plan execution monitoring also plays an
% important role in the GIVE problem.  At a high level, the system needs
% to monitor a user's actions and compare them against the generated
% instruction set to determine if the user has correctly followed
% directions or not. In the case of the latter, new instructions may
% have to be generated. In practice, the situation can be quite
% complicated since the mental state of the user is not known and so the
% system must be inferred from observing the user's actions in real
% time. For example, a user directed to ``turn around and walk through
% the door'' may not necessarily perform these actions to the letter,
% i.e., immediately turning 180 degrees and proceeding directly to the
% door. Instead, the user might take a roundabout route through the
% room, eventually exiting out the door.  Although the user's actions do
% not match the generated instructions exactly, they meet the intended
% goal. The system must be able to identify such ``equivalent'' plans
% and not immediately generate new instructions as soon as the user's
% actions have gone off course. Furthermore, a user can communicate
% certain intentions to the system, both through action and
% inaction. For instance, the system should infer that a user has failed
% to follow instructions if the user exits a room when given a directive
% to ``walk to the centre of the room''. The system should also make a
% similar conclusion if a user simply does nothing when given the
% instruction.

@@

\todo{planning claims general applicability; it's clear that IPC
  benchmarks focus areas of improvement; but it's not so clear to what
  extent this limits usefulness for non-IPC application areas; this
  paper evaluates this (for the first time??)}
